Substructure Boosts to Dark Matter Annihilation from Sommerfeld Enhancement

### 2009-02-17 tex/somm.tex
Added to total and reduced boost section and dSph section

\def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}

@@

\subsection{Some details of the particle physics that do not particularly concern us here}

@@

The contribution of the \DM\ density profile on the structure quantity
\lum\ can be simply evaluated in the context of any density model. As
in Ref.~\cite{2007PhRvD..75h3526S} we assume that the distance
\dist\ is much larger than the scale-radius of the \DM\ halo. For an
\NFW\ profile 90\% of the flux originates from the region with one
scale-radius, and we will, conservatively, only take contributions to
the \DM\ annihilation flux coming rom this region for all the
\GNFW\ profiles. Therefore, we find that 
\begin{equation}\label{eq:lumrhosrsNFW}
\int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd
l\, \rho^2(l)= \normnfw(\gamma)\, \frac{\rhos^2\, \rs^3}{\dist^2}\, .
\end{equation}
The normalization of this relationship depends on the logarithmic
slope of the inner cusp and is given by
\begin{equation}\label{eq:normgammaNFW}
\normnfw(x) = 4 \pi  \, \frac{2^{2x-5}\left(-16+11x-2x^2\right)}{-30+47x-24x^2+4x^3}\, .
\end{equation}
The normalization for the \NFW\ profile is $7\pi/6$ and it depends
strongly on the value of the logarithmic slope of the inner density
profile, as can be seen in Figure \ref{fig:normgammaNFW}. The lower
panel of Fig.~\ref{fig:normgammaNFW} shows the fraction of the \DM\
annihilation flux coming from region within one scale-radius of the
halo. For reasonable values of the inner slope this fraction is in the
range 80 -- 95\%.

@@

Before proceeding we should specify the upper and lower limits of the
integral appearing in \eqnname\ (\ref{eq:boost}). The integration over
ever smaller substructures is cut-off at the size of the smallest
subhalos. The mass scale at which this cut-off occurs is the thermal
free-streaming scale, a \DM\ model-dependent quantity, which in many
\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol
\cite{1999PhRvD..59d3517S,2001PhRvD..64h3507H,2004MNRAS.353L..23G,2005PhRvD..71j3520L,2006PhRvL..97c1301P}. The
upper limit to this integral is some fraction $q$ of the halo mass
$M$, since the abundance of substructure does not continue up to this
mass scale.

@@

\begin{figure*}[t]
\includegraphics[clip=]{2dboost8_65_0.8.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{2dboost8_65_1.4.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{2dboost8_65_-1.0.eps}%%BoundingBox: 190 291 430 474
\caption{The total boost for a $M=10^8$ \Msun\ subhalo assuming a
\GNFW\ dark matter density profile for the subhalos.}%
\label{fig:subboostprofiles}%
\end{figure*}

@@

This is confirmed by looking at the ``reduced boost'' defined as
\begin{equation}\label{eq:redboost}
\redboost(M) \equiv \frac{\totalboost(M)}{\somm(M)\left(1+\boost_0(M)\right)}\, ,
\end{equation}
in which $\boost_0(M)$ is the boost we would find if we turned off the
Sommerfeld enhancement in the calculation of the total boost, \ie, the
boost we would get if we set $\somm(m)$ equal to one in \eqnname\
(\ref{eq:boost}). This $\boost_0$ corresponds to the boost factor that
has been used before to estimate the total boost from density
enhancements in \dSphs\ and it is generally of order \order(10). This
reduced boost therefore divides out the contribution of the Sommerfeld
enhancement of the top-level subhalo and the boost from the
overdensities in lower-level halos, and it would be equal to one if
there were no extra boost coming from the fact that the substructures
are kinematically colder than the main subhalo. As we can see in
Figure \ref{fig:subboost} (\emph{lower panels}), in various parts of
the Yukawa coupling parameter space this reduced boost is equal to one
such that there is no extra boosts due to the lower velocity
dispersion in the substructures. This is mainly the case for large
values of the parameter \mv/\mdm, since the Sommerfeld enhancement, in
the 1/$v$ regime, levels off around $\somm \sim \mv/\mdm$, and since
the top-level subhalo has a velocity dispersion of $\sim 10^{-5}$ this
means that the Sommerfeld enhancement has mostly saturated at these
large values of \mv/\mdm, even in regions of parameter space in which
$\alpha$ is small. Indeed, in the 1/$v$ regime the Sommerfeld
enhancement is $\sim \alpha/\beta$ $\approx$ $10^{-2}$ for $\alpha =
10^{-3}$, such that the Sommerfeld enhancement has fully saturated
over the whole $\alpha$ range for large values of \mv/\mdm\ ($\mv/\mdm
\sim 10^{-2}$).

@@

The total boost factors obtained by assuming an \NFW\ dark matter
density profile do not depend much on the inner slope of the density
profile. This is mainly because the calculation of the boost only
considers \emph{relative} luminosities, for which the difference in
normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
different inner slopes is irrelevant, and because the function $f(c)$
is only a weak function of the inner slope. This is confirmed by
calculating the total boost for a $M = 10^8$ \Msun subhalo for a
subhalo with a very shallow inner cusp as well as for a subhalo with a
very steep inner slope. These are shown in Figure
\ref{fig:subboostprofiles}, and, from comparing this with the total
boost shown in Figure \ref{fig:subboost}, it is clear that the inner
slope does not affect the total boost due to substructure. The main
difference between the different dark matter density profiles is
therefore given by the normalization of the $\lum \propto \rhos^2
\rs^3/\dist^2$ relation, shown in Figure \ref{fig:normgammaNFW}.

BOVY: describe different slopes $n$ here.

@@

@@

\begingroup
%\squeezetable
\begin{table*}[t]
\begin{ruledtabular}
\begin{tabular}{lrr@{.}lr@{}lr@{}lr@{.}lr@{.}lr@{.}l@{ $\times$ }lc}
dSph & \dist\ & \multicolumn{2}{c}{\sigv}\ & \multicolumn{2}{c}{\rmax} & \multicolumn{2}{c}{\vmax} & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{3}{c}{$M(<\rs)$} & References\\
& [kpc] & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[$\Msun$ kpc$^{-3}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{3}{c}{[$\Msun$]}&\\[4pt]
\hline
Carina.......................................................................... & 101 & 6&8 & 3&\phantom{f}      & 16&$\!\!\!$      & 1&1 $\times$ $10^7$ & 1&4 & 7&4&$10^8$ & 1,3\\
Coma Berenices...........................................................      & 44  & 4&6 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&10$^7$ & 2,4\\
Draco........................................................................... & 80  & 5&5 & 8&      & 40&      & 1&0 $\times$ $10^7$ & 3&7 & 1&2&$10^9$ & 1,3\\
Fornax.........................................................................  & 138 & 11&1& 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
Leo I............................................................................  & 250 & 8&8 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
Leo II...........................................................................  & 205 & 6&7 & 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
Sculptor.......................................................................  & 79  & 8&5 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
Sextans........................................................................  & 86  & 5&8 & 2&      & 10&      & 1&0 $\times$ $10^7$ & 0&9 & 1&9&$10^7$ & 1,3\\
Ursa Major II..............................................................   & 32  & 6&7 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&$10^7$ & 2,4\\
Ursa Minor..................................................................    & 66  & 15&0& 6&      & 30&      & 1&0 $\times$ $10^8$ & 2&8 & 5&2&$10^8$ & 1,3\\
%Willman I............. & 38 &&& 5&5 $\times$ $10^8$ & 0&2 & 1&10$^{15}$\\
%Segue I.................. & 23 &&&&
\end{tabular}
\end{ruledtabular}
\caption{Properties of the dwarf Spheroidals used in this study.\\
References: (1) \citet{2008ApJ...672..904P}, (2) \citet{2008ApJ...678..614S}, (3) \citet{1998ARA&A..36..435M}, (4) \citet{2007ApJ...670..313S}.}\label{tab:dsphs}
\end{table*}
\endgroup

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_leoI_65.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{flux_leoII_65.eps}\\%%BoundingBox: 178 291 400 474
%\includegraphics[clip=]{flux_sculptor_65.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{flux_sextans_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
\caption{}%
\label{fig:dsph}%
\end{figure*}

@@

\begin{figure}[t]
\centering
\includegraphics{hist_-2_-3_10_1.eps}
\caption{Distances and masses in \VL\ 2.}%\
\label{fig:VL23_1}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-2_-3_10_2.eps}
\caption{Photon count for individual subhalos and total boosts for \ldots.}%\
\label{fig:VL23_2}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-2_-3_10_3.eps}
\caption{Distribution of total photon counts for 1,000 random observers at 8.5 kpc.}%\
\label{fig:VL23_3}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-5_-1_10_2.eps}
\caption{}%\
\label{fig:VL51_1}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-5_-1_10_3.eps}
\caption{}%\
\label{fig:VL51_2}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-3.84375_-2.625_10_2.eps}
\caption{}%\
\label{fig:VLresonance1}%
\end{figure}

\begin{figure}[t]
\centering
\includegraphics{hist_-3.84375_-2.625_10_3.eps}
\caption{}%\
\label{fig:VLresonance2}%
\end{figure}

@@

{\bf Acknowledgments:}
Hogg. NASA NNX08AJ48G.

>>>>

%Some voodoo
\def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}
\renewcommand{\topfraction}{0.95}
\renewcommand{\bottomfraction}{0.95}
\renewcommand{\dbltopfraction}{0.95}

@@

\subsection{Some details of the particle physics that do not particularly concern us here}\label{sec:partphys}

@@

The contribution of the \DM\ density profile on the structure quantity
\lum\ can be simply evaluated in the context of any density model. As
in Ref.~\cite{2007PhRvD..75h3526S} we assume that the distance
\dist\ is much larger than the scale-radius of the \DM\ halo. For an
\NFW\ profile 90\% of the flux originates from the region with one
scale-radius, and we will, conservatively, only take contributions to
the \DM\ annihilation flux coming rom this region for all the
\GNFW\ profiles. Therefore, we find that 
\begin{equation}\label{eq:lumrhosrsNFW}
\int_{\Delta\Omega}\,\frac{\dd\Omega}{4\pi}\,\int_{\mathrm{\los}} \dd
l\, \rho^2(l)= \normnfw(\gamma)\, \frac{\rhos^2\, \rs^3}{\dist^2}\, .
\end{equation}
The normalization of this relationship depends on the logarithmic
slope of the inner cusp and is given by
\begin{equation}\label{eq:normgammaNFW}
\normnfw(x) = 4 \pi  \, \frac{2^{2x-5}\left(-16+11x-2x^2\right)}{-30+47x-24x^2+4x^3}\, .
\end{equation}
The normalization for the \NFW\ profile is $7\pi/6$ and it depends
strongly on the value of the logarithmic slope of the inner density
profile, as can be seen in Figure \ref{fig:normgammaNFW}. The bottom
panel of Fig.~\ref{fig:normgammaNFW} shows the fraction of the \DM\
annihilation flux coming from region within one scale-radius of the
halo. For reasonable values of the inner slope this fraction is in the
range 80 -- 95\%.

@@

Before proceeding we should specify the upper and lower limits of the
integral appearing in \eqnname\ (\ref{eq:boost}). The integration over
ever smaller substructures is cut-off at the size of the smallest
subhalos. The mass scale at which this cut-off occurs is the thermal
free-streaming scale, a \DM\ model-dependent quantity, which in many
\DM\ models lies in the range 10$^{-6}$ -- 10$^{-12}$ \Msol\
\cite{1999PhRvD..59d3517S,2001PhRvD..64h3507H,2004MNRAS.353L..23G,2005PhRvD..71j3520L,2006PhRvL..97c1301P}. The
upper limit to this integral is some fraction $q$ of the halo mass
$M$, since the abundance of substructure does not continue up to this
mass scale.

@@

\begin{figure*}[t]
\includegraphics[clip=]{2dboost8_65_0.8.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{2dboost8_65_1.4.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{2dboost8_65_0.1.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{2dboost8_65_0.2.eps}%%BoundingBox: 190 291 430 474
\caption{The total boost for a $M=10^8$ \Msun\ subhalo for different shapes of the \DM\ density profile: a generalized \NFW\ profile described by an inner slope $\gamma$ (\emph{top panels}); an Einasto profile described by a shape parameter \alphaEinasto\ (\emph{bottom panels}).}%
\label{fig:subboostprofiles}%
\end{figure*}

@@

This is confirmed by looking at the ``reduced boost'' defined as
\begin{equation}\label{eq:redboost}
\redboost(M) \equiv \frac{\totalboost(M)}{\somm(M)\left(1+\boost_0(M)\right)}\, ,
\end{equation}
in which $\boost_0(M)$ is the boost we would find if we turned off the
Sommerfeld enhancement in the calculation of the total boost, \ie, the
boost we would get if we set $\somm(m)$ equal to one in \eqnname\
(\ref{eq:boost}). This $\boost_0$ corresponds to the boost factor that
has been used before to estimate the total boost from density
enhancements in \dSphs\ and it is generally of order \order(10). This
reduced boost therefore divides out the contribution of the Sommerfeld
enhancement of the top-level subhalo and the boost from the
overdensities in lower-level halos, and it would be equal to one if
there were no extra boost coming from the fact that the substructures
are kinematically colder than the main subhalo. As we can see in
Figure \ref{fig:subboost} (\emph{bottom panels}), in various parts of
the Yukawa coupling parameter space this reduced boost is equal to one
such that there is no extra boosts due to the lower velocity
dispersion in the substructures. This is mainly the case for large
values of the parameter \mv/\mdm, since the Sommerfeld enhancement, in
the 1/$v$ regime, levels off around $\somm \sim \mv/\mdm$, and since
the top-level subhalo has a velocity dispersion of $\sim 10^{-5}$ this
means that the Sommerfeld enhancement has mostly saturated at these
large values of \mv/\mdm, even in regions of parameter space in which
$\alpha$ is small. Indeed, in the 1/$v$ regime the Sommerfeld
enhancement is $\sim \alpha/\beta$ $\approx$ $10^{-2}$ for $\alpha =
10^{-3}$, such that the Sommerfeld enhancement has fully saturated
over the whole $\alpha$ range for large values of \mv/\mdm\ ($\mv/\mdm
\sim 10^{-2}$).

\begin{figure*}[t]
\includegraphics[clip=]{2dboost8_65_-0.7.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{2dboost8_65_-1.0.eps}\\%%BoundingBox: 190 291 430 474
%\includegraphics[clip=]{2dboost8_65_-3.eps}%%BoundingBox: 178 291 400 474
%\includegraphics[clip=]{2dboost8_65_-12.eps}%%BoundingBox: 190 291 430 474
\caption{The total boost for a $M=10^8$ \Msun\ for different slopes
$n$ for the subhalo mass function (\emph{top panels}) and for
different values of the dark matter thermal free streaming limit
$m_0$ (\emph{bottom panels}).}%
\label{fig:subboostslope}%
\end{figure*}

@@

The total boost factors obtained by assuming an \NFW\ dark matter
density profile do not depend much on the inner slope of the density
profile. This is mainly because the calculation of the boost only
considers \emph{relative} luminosities, for which the difference in
normalization of the $\lum \propto \rhos^2 \rs^3/\dist^2$ relation for
different inner slopes is irrelevant, and because the mass of the
subhalo is only weakly affected by the value of the inner slope ---the
function $f(c)$ is only a weak function of the inner slope. This is
confirmed by calculating the total boost for a $M = 10^8$ \Msun\
subhalo for a subhalo with a very shallow inner cusp as well as for a
subhalo with a very steep inner slope. These are shown in Figure
\ref{fig:subboostprofiles} (\emph{top panels}), and, from comparing
this with the total boost shown in Figure \ref{fig:subboost}, it is
clear that the inner slope does not affect the total boost due to
substructure. The main difference between the different dark matter
density profiles is therefore given by the normalization of the $\lum
\propto \rhos^2 \rs^3/\dist^2$ relation, shown in Figure
\ref{fig:normgammaNFW}.

If the \DM\ density profiles of halos is better described by an
Einasto profile then the substructure boost is affected more strongly
by the shape parameter \alphaEinasto\, which basically describes how
cored the \DM\ density profile is. Changing \alphaEinasto\ has a
greater impact on changing the overall \DM\ density profileand
therefore also plays a larger role in the relation between the
characteristic density, the concentration, and the mass of the
subhalo. The total boosts for a $M$ = $10^8$ \Msun\ subhalo is shown
in Figure \ref{fig:subboostprofiles} (\emph{bottom panels}) for two
different values of the shape parameter \alphaEinasto\, which span the
range found in numerical simulations (for values of \alphaEinasto\
greater than 0.2 the behavior basically saturates at that of
\alphaEinasto\ = 0.2). For \alphaEinasto\ = 0.1 the total boost is
about the same as for the \GNFW\ profile shape, which is
understandable given that the Einasto profile is more cuspy for
smaller values of \alphaEinasto. The same dependence on the parameter
\mv/\mdm\ as is seen for the \GNFW\ profile is displayed for the
Einasto profile. In the case that \alphaEinasto\ is larger the
substructure boost reaches slightly larger values than for the \GNFW\
density profile. Thus, as was the case for the \GNFW\ profile, the
impact of the different normalization of the
$\rhominustwo^2\rminustwo^3/\dist^2$ for different values of
\alphaEinasto\ will play a leading role in the difference between the
different profile shapes.

\begingroup
%\squeezetable
\begin{table*}
\begin{ruledtabular}
\begin{tabular}{lrr@{.}lr@{}lr@{}lr@{.}lr@{.}lr@{.}l@{ $\times$ }lc}
dSph & \dist\ & \multicolumn{2}{c}{\sigv}\ & \multicolumn{2}{c}{\rmax} & \multicolumn{2}{c}{\vmax} & \multicolumn{2}{c}{\rhos} & \multicolumn{2}{c}{\rs} & \multicolumn{3}{c}{$M(<\rs)$} & References\\
& [kpc] & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{2}{c}{[km s$^{-1}$]} & \multicolumn{2}{c}{[$\Msun$ kpc$^{-3}$]} & \multicolumn{2}{c}{[kpc]} & \multicolumn{3}{c}{[$\Msun$]}&\\[4pt]
\hline
Carina.......................................................................... & 101 & 6&8 & 3&\phantom{f}      & 16&$\!\!\!$      & 1&1 $\times$ $10^7$ & 1&4 & 7&4&$10^8$ & 1,3\\
Coma Berenices...........................................................      & 44  & 4&6 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&10$^7$ & 2,4\\
Draco........................................................................... & 80  & 5&5 & 8&      & 40&      & 1&0 $\times$ $10^7$ & 3&7 & 1&2&$10^9$ & 1,3\\
Fornax.........................................................................  & 138 & 11&1& 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
Leo I............................................................................  & 250 & 8&8 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
Leo II...........................................................................  & 205 & 6&7 & 4&      & 20&      & 1&0 $\times$ $10^7$ & 1&9 & 1&5&$10^8$ & 1,3\\
Sculptor.......................................................................  & 79  & 8&5 & 6&      & 30&      & 1&0 $\times$ $10^7$ & 2&8 & 5&2&$10^8$ & 1,3\\
Sextans........................................................................  & 86  & 5&8 & 2&      & 10&      & 1&0 $\times$ $10^7$ & 0&9 & 1&9&$10^7$ & 1,3\\
Ursa Major II..............................................................   & 32  & 6&7 & $\cdot$&& $\cdot$& & 3&0 $\times$ $10^8$ & 0&3 & 2&0&$10^7$ & 2,4\\
Ursa Minor..................................................................    & 66  & 15&0& 6&      & 30&      & 1&0 $\times$ $10^8$ & 2&8 & 5&2&$10^8$ & 1,3\\
%Willman I............. & 38 &&& 5&5 $\times$ $10^8$ & 0&2 & 1&10$^{15}$\\
%Segue I.................. & 23 &&&&
\end{tabular}
\end{ruledtabular}
\caption{Properties of the dwarf Spheroidals used in this study.\\
References: (1) \citet{2008ApJ...672..904P}, (2) \citet{2008ApJ...678..614S}, (3) \citet{1998ARA&A..36..435M}, (4) \citet{2007ApJ...670..313S}.}\label{tab:dsphs}
\end{table*}
\endgroup

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_leoI_65.eps}\\%%BoundingBox: 190 291 430 474
\includegraphics[clip=]{flux_leoII_65.eps}%%BoundingBox: 178 291 400 474
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 291 430 474
\caption{}%
\label{fig:dsph}%
\end{figure*}

In Figure \ref{fig:subboostslope} the effect of varying two more of
the many parameters introduced in the previous section is shown. The
slope of the subhalo mass function $n$ was set to -0.9 before,
corresponding to the results from the latest high-resolution numerical
simulations. In Figure \ref{fig:subboostslope} (\emph{top panels}) we
show the effect of both a shallower as well as a steeper subhalo mass
function. The value of $n = -0.7$ corresponds to the shallow slope
found in Ref.~\cite{2002PhRvD..66f3502H}. We see that the effect of
such a shallow slope is to seriously reduce the substructure boost,
which is simply due to the fact that for a shallower slope less
substructure is found in smaller and more concentrated clumps. A
steeper slope such as $n = -1$, as was found in the recent \VL\
simulation \cite{2008ApJ...686..262K} leads to a substructure boost
which is about an order of magnitude larger than for the fiducial
$n=-0.9$ value.

The effect of the thermal free-streaming scale of \DM\ on the
substructure boost is shown in Figure \ref{fig:subboostslope}
(\emph{bottom panels}). This effect is relatively large since the
free-streaming scale sets the scale of the smallest and most
concentrated \DM\ halos and raising or lowering this scale therefore
changes an important part of the contribution to the substructure
boost. For very large values of the free-streaming scale such as $m_0
= 10^{-3}$ \Msun, the substructure boost becomes relatively small
BOVY: BE MORE PRECISE HERE. When the free-streaming scale becomes very
small, much of the substructure is in very small, cold clumps, such
that the total boost becomes very large. BOVY: BE MORE PRECISE HERE.

@@

In order to calculate the substructure boost to the \DM\ annihilation
cross section we need the mass of the \dSphs. Many definitions of the
mass of the galaxy could be used here, among the more appropriate ones
is the mass contained inside the tidal radius, \ie, the radius inside
of which material in the \dSph\ is safe from being tidally stripped by
the Galaxy. However, since we are considering the \DM\ annihilation
flux coming from within the radius \rminustwo\ we will use the mass
contained within this radius to calculate the substructure boost. This
is more appropriate since (1) it can be calculated straightforwardly
from the \DM\ density profile of the \dSph\ and (2) the mass contained
within \rminustwo\ is really the more relevant mass to calculate the
substructure boost, since we are only interested in contributions to
the \DM\ annihilation flux coming from within the radius \rminustwo.

Because of the similarities between some of the \DM\ density profiles
of the \dSphs\ in Table \ref{tab:dsphs} we do not show the predicted
annihilation fluxes for all of the \dSphs\ in Figure \ref{fig:dsph},
instead focusing on a representative sample of them. Inspection of
Table \ref{tab:dsphs} shows that annihilation flux of the \dSphs\ not
shown here will be similar to that of the ones shown in Figure
\ref{fig:dsph}.

It is clear from Figure \ref{fig:dsph} that the \DM\ annihilation flux
predicted to be coming from the \dSphs\ is very large in a large class
of models. For example, the flux predicted for Draco exceeds the upper
limit set on the \DM\ annihilation flux set by
\cite{2008ApJ...679..428A} for all models parametrized by \mv/\mdm\
and $\alpha$ by many orders of magnitude. We have to stress here that
our predicted fluxes are all for the particle physics specific
parameters given in Section \ref{sec:partphys}, \ie, \sigmaannv\ = $3
\times 10^{-26}$ cm$^3$ s$^{-1}$, \mdm\ = 700 GeV and a particular
form for the \DM\ annihilation spectrum. To predict the gamma-ray flux
for a particular model with different values of these particle physics
parameters one is required to calculate the particle physics parameter
specific factor to the \DM\ annihilation flux and appropriately
rescale the values given here. Nevertheless, any model that predicts
very large annihilation cross sections to account for the PAMELA/ATIC
excesses, and we are thinking here specifically about models that try
to fit the PAMELA data with very large boost factors with respect to
the fiducial annihilation cross section, give rise to these very large
predicted gamma-ray fluxes coming from the \dSphs. Therefore, unless
the boosts from substructure are somehow suppressed or the assumptions
about the subhalo mass function, concentration-mass relation for
subhalos, and/or any of the other asumptions about subhalos beyond the
resolution limits of the state-of-the-art numerical simulations are
violated for physical subhalos, these models are in tension with the
experimental limits on the gamma-ray flux from \dSphs.

@@

@@

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_1.eps}
\caption{The distances to and masses of subhalos within the tidal
radius of the main halo in \VL\ 2. The distances are for three random
observers at a distance of 8.5 kpc from the galactic center of the
main halo.}%\
\label{fig:VL23_1}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_2.eps}
\caption{Left: Histogram of the predicted gamma-ray flux coming from
indivual subhalos contained within the tidal radius of the main halo
in \VL\ 2 for three random observer placed at a distance of 8.5 kpc
from the center of the main \VL\ halo for \mv/\mdm\ = 10$^{-2}$ and
$\alpha = 10^{-3}$. Right: Histogram of the total boosts of the
subhalos for the same values of \mv/\mdm\ and $\alpha$.}%\
\label{fig:VL23_2}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-2_-3_10_3.eps}
\caption{Total gamma-ray flux from subhalos in \VL\ 2 for 1,000 random
observers at 8.5 kpc from the center of the main \VL\ halo for
\mv/\mdm\ = 10$^{-2}$ and $\alpha = 10^{-3}$.}%\
\label{fig:VL23_3}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-5_-1_10_2.eps}
\caption{Same as Figure \ref{fig:VL23_2} but for \mv/\mdm\ = 10$^{-5}$
and $\alpha = 10^{-1}$.}%
\label{fig:VL51_1}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-5_-1_10_3.eps}
\caption{Same as Figure \ref{fig:VL23_3} but for \mv/\mdm\ = 10$^{-5}$
and $\alpha = 10^{-1}$.}%
\label{fig:VL51_2}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-3.84375_-2.625_10_2.eps}
\caption{Same as Figure \ref{fig:VL23_2} but for a resonance:
\mv/\mdm\ = 10$^{-3.84375}$ and $\alpha = 10^{-2.625}$.}%
\label{fig:VLresonance1}%
\end{figure}

\begin{figure}
\centering
\includegraphics{hist_-3.84375_-2.625_10_3.eps}
\caption{Same as Figure \ref{fig:VL23_3} but for a resonance:
\mv/\mdm\ = 10$^{-3.84375}$ and $\alpha = 10^{-2.625}$.}%
\label{fig:VLresonance2}%
\end{figure}

@@

{\bf Acknowledgments:} It is a pleasure to thank Mulin Ding and David
Hogg for comments and assistance. This research was partially
supported by NASA (grant NNX08AJ48G).
